First Season MWA EoR Power Spectrum Results at Redshift 7

### 2016-06-27 ms.tex
BJH comments. Have question out to Mahsa re: peeling bug. Will update fig. 13 with MFM comments

While the 21 cm EoR signal is priority, the MWA is a general observatory serving several 
science programs beyond cosmology including Galactic and extragalactic surveys, time 
domain astrophysics, solar monitoring, and ionospheric science. The array was thus 
designed with these programs in consideration and the layout was optimized using a 
pseudo-random antenna placement algorithm \citep{Beardsley:2012} to obtain a well 
behaved point spread function while retaining a dense core for EoR sensitivity 
\citep{Beardsley:2013}. This is in contrast to more pointed EoR experiments like PAPER, 
which utilize a highly redundant layout to enhance sensitivity. Redundant arrays can 
leverage symmetries of the instrument for quick calibration and analysis, but have poor 
point spread functions, making foregrounds more difficult to characterize. Conversely we 
use the high imaging capability of the MWA to calibrate and subtract foregrounds based on 
sky models. A full description of the science capabilities of the array is found in 
\citealt{Bowman:2013}.

@@

Our final data cut was made when inspecting the residual continuum images output from 
FHD. We calculate a fractional residual flux RMS in the following way. For each snapshot 
continuum image, we find the residual flux in the pixels of all sources greater than 0.5 Jy 
subtracted within half-beam power and scale by the integrated flux of the sources. This 
gives us a list of fractional residual fluxes for the snapshot, which we then use to calculate 
an RMS. We found that most snapshots had residual flux RMS $<10\%$ with some outliers, 
which we cut from the data. This cut largely overlapped with previous cuts, but removed an 
additional 95 observations. The cause of the high residual flux RMS is yet unknown and is 
not localized by observation day, LST, pointing, or any other jackknife we have performed.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{slices.pdf}
\caption{
\emph{Left:} The N-S, $z=6.8$ two dimensional power spectrum repeated from 
Figure~\ref{fig:2d_ps_subband}. Here we have superimposed blue lines to show slices in
$k_{\perp}$ (solid) and $k_{||}$ (dashed). \emph{Top right:} Residual power as a function
of $k_{\perp}$ for fixed $k_{||}$. The grey boxes show regions which will be excluded from
the one dimensional averages. The red dashed line shows the 1-$\sigma$ noise level, and
the vertical black dashed line shows the intersection of this slice with the wedge. 
\emph{Bottom right:} Residual power as a function of $k_{||}$ for fixed $k_{\perp}$. The strong
foreground contamination is evident at low $k_{||}$, and the coarse band harmonic lines
appear as expected. The thin black line in both right panels is the equivalent power slice from the RTS$+$CHIPS comparison pipeline, which is discussed in Section~\ref{subsec:ref_pipe}.
\label{fig:slices}
}
\end{center}
\end{figure*}

@@

Finally, we place our best upper limits in context with the 21cm EoR field as a whole. A 
direct comparison between measurements from different instruments is difficult due to the 
varying methods, redshifts, and scales probed. However, if we approximate the power 
spectrum in units of mK$^2$ to be roughly flat over the scales probed by the MWA, PAPER, 
and GMRT, we can glean a view of the current state of the field. The results from this 
analysis are a modest improvement over the previous best MWA results \citep{Dillon:2015}, 
which is not surprising given that this integration contains about ten times more data, but 
has a strong signal to noise on the power spectrum. Our results are also consistent with the 
GMRT limit of $\Delta^2\le6.15\times10^4$~mK$^2$ at $z=8.6$ \citep{Paciga:2013}, though 
probe significantly different periods of the EoR. The frontrunner in terms of an upper limit is 
the PAPER experiment \citep{Ali:2015, Jacobs:2015, Parsons:2014}. The PAPER-64 limit of 
$\Delta^2\le502$~mK$^2$ at $z=8.4$ is about one and a half orders of magnitude lower 
than our best limit in power units, though again probing a significantly different redshift. 
While the MWA analysis is not yet at that level of maturity, there is room for rapid 
improvement. The dashed red line in Figure~\ref{fig:1d_ps} indicate the noise level possible 
to reach with this data set if systematics can be overcome.

>>>>

While the 21 cm EoR signal is priority, the MWA is a general observatory serving several 
science programs beyond cosmology including Galactic and extragalactic surveys, time 
domain astrophysics, solar monitoring, and ionospheric science. The array was thus 
designed with these programs in consideration and the layout was optimized using a 
pseudo-random antenna placement algorithm \citep{Beardsley:2012} to obtain a well 
behaved point spread function while retaining a dense core for EoR sensitivity 
\citep{Beardsley:2013}. 
A full description of the science capabilities of the array is found in
\citealt{Bowman:2013}.
The high imaging capability of the MWA enables us to calibrate and subtract foregrounds
based on sky models, leveraging the full capability of the observatory for the
EoR experiment. 
This strategy is in contrast to more pointed EoR experiments like PAPER, 
which utilize a highly redundant layout to enhance sensitivity. Redundant arrays can 
exploit symmetries of the instrument for quick calibration and analysis, but have poor 
point spread functions, making foregrounds more difficult to characterize.

@@

Our final data cut was made when inspecting the residual continuum images output from 
FHD. We test for residual flux at source locations by calculating a fractional residual flux RMS 
in the following way. For each snapshot 
continuum image, we find the residual flux in the pixels of all sources greater than 0.5 Jy 
subtracted within half-beam power and scale by the integrated flux of the sources. This 
gives us a list of fractional residual fluxes for the snapshot, which we then use to calculate 
an RMS. We found that most snapshots had residual flux RMS $<10\%$ with some outliers, 
which we cut from the data. This cut largely overlapped with previous cuts, but removed an 
additional 95 observations. The cause of the high residual flux RMS is yet unknown and is 
not localized by observation day, LST, pointing, or any other jackknife we have performed.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{slices.pdf}
\caption{
\emph{Left:} The N-S, $z=6.8$ two dimensional power spectrum repeated from 
Figure~\ref{fig:2d_ps_subband}. Here we have superimposed blue lines to show slices in
$k_{\perp}$ (solid) and $k_{||}$ (dashed). \emph{Top right:} Residual power as a function
of $k_{\perp}$ for fixed $k_{||}$, averaged in rings from the full 3D power cube (blue). The grey boxes show regions which will be excluded from
the one dimensional averages. The red line shows the 1-$\sigma$ noise level, and
the vertical black dashed line shows the intersection of this slice with the wedge. 
\emph{Bottom right:} Residual power as a function of $k_{||}$ for fixed $k_{\perp}$, again averaged in rings from full 3D power cube (blue). The strong
foreground contamination is evident at low $k_{||}$, and the coarse band harmonic lines
appear as expected. The thin black line in both right panels is the equivalent power slice from the RTS$+$CHIPS comparison pipeline, which is discussed in Section~\ref{subsec:ref_pipe}.
\label{fig:slices}
}
\end{center}
\end{figure*}

@@

Finally, we place our best upper limits in context with the 21cm EoR field as a whole. A 
direct comparison between measurements from different instruments is difficult due to the 
varying methods, redshifts, and scales probed. However, if we approximate the power 
spectrum in units of mK$^2$ to be roughly flat over the scales probed by the MWA, PAPER, 
and GMRT, we can glean a view of the current state of the field. The results from this 
analysis are a modest improvement over the previous best MWA results \citep{Dillon:2015}, 
which is not surprising given that this integration contains about ten times more data, but 
has a strong signal to noise on the power spectrum. Our results are also consistent with the 
GMRT limit of $\Delta^2\le6.15\times10^4$~mK$^2$ at $z=8.6$ \citep{Paciga:2013}, though 
probe significantly different periods of the EoR. The frontrunner in terms of an upper limit is 
the PAPER experiment \citep{Ali:2015, Jacobs:2015, Parsons:2014}. The PAPER-64 limit of 
$\Delta^2\le502$~mK$^2$ at $z=8.4$ is about one and a half orders of magnitude lower 
than our best limit in power units, though again probing a significantly different redshift. 
While the MWA analysis is not yet at that level of maturity, there is room for rapid 
improvement. The red lines in Figure~\ref{fig:1d_ps} indicate the noise level possible 
to reach with this data set if systematics can be overcome.
